{\fontfamily{ptm}\selectfont The Proximal Point interpretation of Nesterov accelerated proximal gradient}

### 2024-08-15 Notebook LaTEX/Momentum Accelerations/paper.tex
TwT, a lot of changes

\author{
    Author 1 Name, Author 2 Name
    \thanks{
        Subject type, Some Department of Some University, Location of the University,
        Country. E-mail: \texttt{author.name@university.edu}.
    }
}

@@

\section{Generic Lyapunov analysis for Accelerated gradient via PPM}
\label{sec:generic_ag_ppm_lyapunov_analysis}
    The following lemma establish an upper bound for quantities used in the Lypunov functions for the convergence rate of 
    \hyperref[def:ag_prox_grad_generic]
    {definition \ref*{def:ag_prox_grad_generic}}
    \begin{lemma}[Lyapunov Inequality]. 
    The Lypunov function in use is identical to Ahn and Sra \cite[section 4.2]{ahn_understanding_2022} except for the involvement of a gradient mapping. 
    We present the results differently by compatifying results into 
    \hyperref[lemma:nsmooth_agg_lyapunov_upper_bound]
    {lemma \ref*{lemma:nsmooth_agg_lyapunov_upper_bound}}. 
    Theorem \ref*{thm:generic_ag_convergence} 
    states new results that are extension to Ahn and Sra's works by introducing an error term to the Lypunov analysis, allowing for a weaker constraints for the stepsize paramters $\tilde \eta_i, \eta_i$ from 
    \hyperref[def:ag_prox_grad_generic]{definition \ref*{def:ag_prox_grad_generic}}. 
    \par
    Theorem \ref*{thm:ag_generic_stepsize_constrants} states the weakened constraints on the stepsize parameters $\tilde \eta_i, \eta_i$.

@@

        Let $h = f + g$ be a sum of $f$ that is $L$-Lipschitz smooth convex, and $g$ be convex.

@@

        This lemma is designed to work with 
        \hyperref[def:ag_prox_grad_generic]{definition \ref*{def:ag_prox_grad_generic}}.

@@

    \begin{theorem}[Generic AG Convergence]
    \label{thm:generic_ag_convergence}
        Define the Lyapunov function $\Phi_t\; \forall t \in \{0\}\cup \N$: 
        \begin{align*}
            \Phi_t &= \left(
                \sum_{i = 1}^{t} \tilde\eta_{i}
            \right) (f(z_t) - f(x_*)) + \frac{1}{2}\Vert x_t - x_*\Vert^2 \quad \forall t \in \NN
            \\
            \Phi_0 &= \frac{1}{2}\Vert x_0 - x_*\Vert^2, 
        \end{align*}
        if there is a choice of $\eta_i, \tilde \eta_i$ in 
        \hyperref[def:ag_prox_grad_ppm]{definition \ref*{def:ag_prox_grad_ppm}} such that we have 
        \begin{align*}
            & \Phi_{t + 1} - \Phi_{t} =
            \left(
                \sum_{i = 1}^{t} \tilde \eta_i
            \right)\Upsilon_{2, t + 1}^{\text{AG}} + 
            \Upsilon_{1, t + 1}^{\text{AG}} 
            \le \delta_{t + 1} \quad 
            \forall t \in \mathbb N, 
            \\
            & \Upsilon_{1, 1}^{\text{AG}} \le \delta_1. 
        \end{align*}
        Then with
        \begin{align*}
            S_t := \sum_{i = 1}^{t} \delta_i,
            \quad 
            \sigma_t := \sum_{i = 1}^{t}\tilde \eta_i, 
        \end{align*}
        we have $\Phi_{t + 1} - \Phi_t \le \delta_{t + 1}$, it allows for a convergence rate of $\mathcal O \left(\sigma_T^{-1}S_T\right)$ of $f(z_T) - f(x_*)$ when $S_T \neq 0$ and a rate of $\mathcal O(\sigma_T^{-1})$ when $S_T \le 0$. 
    \end{theorem}
    \begin{proof}
        By definition we have
        {\footnotesize
        \begin{align*}
            \Phi_{t + 1} - \Phi_t 
            &= 
            \left(
                \sum_{i = 1}^{t+1} \tilde\eta_{i}
            \right) (f(z_{t + 1}) - f(x_*)) 
            - 
            \left(
                \sum_{i = 1}^{t} \tilde\eta_{i}
            \right) (f(z_{t}) - f(x_*)) 
            + \frac{1}{2}\Vert x_t - x_*\Vert^2
            - \frac{1}{2}\Vert x_{t + 1} - x_*\Vert^2
            \\
            &= 
            \tilde \eta_{t + 1} (f(z_{t + 1}) - f(z_*))
            +
            \left(
                \sum_{i = 1}^{t} \tilde \eta_i
            \right)(f(z_{t + 1}) - f(z_t))
            + \frac{1}{2}\Vert x_t - x_*\Vert^2
            - \frac{1}{2}\Vert x_{t + 1} - x_*\Vert^2
            \\
            &= \left(
                \sum_{i = 1}^{t} \tilde \eta_i
            \right)\Upsilon_{2, t + 1}^{\text{AG}} + \Upsilon_{1, t + 1}^{\text{AG}} \le \delta_{t + 1}. 
        \end{align*}
        }
        Telescoping we have 
        \begin{align*}
            \Phi_{t + 1} - \Phi_t &\le \delta_{t + 1}
            \\
            \sum_{i = 0}^{T - 1}\Phi_{i + 1} - \Phi_i &\le \sum_{i = 0}^{T - 1}\delta_{i + 1}
            \\
            \Phi_T - \Phi_0 &\le 
            \sum_{i = 1}^{T}\delta_i = S_{T}. 
        \end{align*}
        So then $\Phi_T - \Phi_0$ yields: 
        $$
        \begin{aligned}
            \sigma_T (f(z_t) - f(x_*)) 
            + \frac{1}{2}\Vert x_t - x_*\Vert^2 
            - \frac{1}{2}\Vert x_0 - x_*\Vert^2 
            &\le S_{T}
            \\
            \implies 
            \sigma_T(f(z_t) - f(x_*))
            &\le 
            S_T + \frac{1}{2}\Vert x_0 - x_*\Vert^2
            \\
            f(z_t) - f(x_*) &\le 
            \sigma_T^{-1}\left(
                S_{T} + \frac{1}{2}\Vert x_0 - x_*\Vert^2
            \right),
        \end{aligned}
        $$
        which yields a convergence rate $\mathcal O(\sigma_T^{-1}S_{T})$. 
        When $S_T = 0$, the convergence rate is $O(\sigma_T^{-1})$ instead. 
    \end{proof}

    \begin{theorem}[Constraints on the PPM stepsize sequence]
        \label{thm:ag_generic_stepsize_constrants}
        \;\\
        Algorithm formulated in 
        converges if the stepsize sequence $\eta_i, \tilde \eta_{i}$ satisfies

@@

        Here $\epsilon_{t + 1} = \tilde \eta_{t + 1} - \eta_t - L^{-1}$ is a sequence made to characterizing the error between $\tilde \eta_{t + 1}, \eta_t - L^{-1}$. 
        Then the convergence rate can be formulated by

@@

        When $\epsilon_t = 0 \;\forall t\in \mathbb N$, the sequence $a_t:= a_t = 1 + L \eta_t = \tilde \eta_{t + 1} - \epsilon_{t + 1}$, the relation simplies to: 
                c_{t + 1} = a_t^2.     
        As a consequence: 
            \item It makes $\delta_{t} = 0$ in the Lypunov analysis; 
            \item it simplifies the algorithm to FISTA. 
            \item and $a_t$ is the famous Nesterov momentum sequence. 

@@

\section{Recovering existing variants of Nesterov accelerated gradient}\label{sec:recovery}
    \subsection{Recovering FISTA}
        From 
        \hyperref[thm:ag_generic_stepsize_constrants]{theorem \ref*{thm:ag_generic_stepsize_constrants}}, 
        setting $\epsilon_i = 0$ for all $i$ yields $a_t = 1 + L\eta_t = \tilde \eta_{t + 1}$, for all $t \in \N$ where $a_t$ is the Nesterov sequence.
        In this case, there exists a momentum form of the generic form and it reduces to FISTA from Beck and Toubolle, or Similar Triangle Method as referred to by Ahn and Sra.

@@

            \hyperref[thm:ag_generic_stepsize_constrants]
                {theorem \ref*{thm:ag_generic_stepsize_constrants}}

>>>>

\author{
    Author 1 Name, Author 2 Name
    \thanks{
        Subject type, Some Department of Some University, Location of the University,
        Country. E-mail: \texttt{alto@mail.ubc.ca}.
    }
}

@@

\section{Generic Lyapunov analysis for Accelerated gradient via PPM}
\label{sec:generic_ag_ppm_lyapunov_analysis}
    In this section we derive the generic convergence rate formuated by $\eta_i, \tilde \eta_i$. 
    \par
    The Lypunov function in 
    \hyperref[thm:generic_ag_convergence]{\ref*{thm:generic_ag_convergence}}
    is identical to Ahn and Sra 
    \cite[section 4.2]{ahn_understanding_2022}, 
    except for the involvement of the gradient mapping. 
    We adapated it into the context of proximal gradient by 
    \hyperref[lemma:nsmooth_agg_lyapunov_upper_bound]
    {lemma \ref*{lemma:nsmooth_agg_lyapunov_upper_bound}}
    \par
    We present the convergence rate by compatifying results into 
    \hyperref[lemma:nsmooth_agg_lyapunov_upper_bound]
    {lemma \ref*{lemma:nsmooth_agg_lyapunov_upper_bound}} 
    to assist the proof of  
    \hyperref[thm:generic_ag_convergence]
    {Theorem \ref*{thm:generic_ag_convergence}} 
    which states new results that are extension to Ahn and Sra's works by introducing an error term to the Lypunov analysis, allowing for a weaker constraints for the stepsize paramters $\tilde \eta_i, \eta_i$ from 
    \hyperref[def:ag_prox_grad_generic]{definition \ref*{def:ag_prox_grad_generic}}. 
    \hyperref[thm:ag_generic_stepsize_constraints]{Theorem \ref*{thm:ag_generic_stepsize_constraints}}
    states the weakened constraints on the stepsize parameters $\tilde \eta_i, \eta_i$. 
    \par
    Through out this section, we continue the assumption for $h = f + g$ and gradient mapping $\mathcal G_L$ the same as previous section.

    \begin{lemma}[Lyapunov Inequality]. 

@@

@@

    \label{remark:upsilon-upperbound-for-iterates}
        Let $y_t, x_t$ be given by
        \hyperref[def:ag_prox_grad_generic]
        {definition \ref*{def:ag_prox_grad_generic}}.

@@

    \begin{theorem}[Generic AG Convergence]
    \label{thm:generic_ag_convergence}
        Let $\Upsilon_{1, t + 1}^\text{AG}, \Upsilon_{2, t + 1}^\text{AG}$ be given by 
        \hyperref[remark:upsilon-upperbound-for-iterates]
        {remark \ref*{remark:upsilon-upperbound-for-iterates}}. 
        Let $z_t$ be given by 
        \hyperref[def:ag_prox_grad_generic]
        {definition \ref*{def:ag_prox_grad_generic}}
        Define the Lyapunov function $\Phi_t\; \forall t \in \{0\}\cup \N$, $S_t$ and $\sigma_t$: 
            \Phi_t &:= \left(
                \sum_{i = 1}^{t} \tilde\eta_{i}
            \right) (h(z_t) - h(x_*)) + \frac{1}{2}\Vert x_t - x_*\Vert^2 \quad \forall t \in \N
            \Phi_0 &:= \frac{1}{2}\Vert x_0 - x_*\Vert^2, 
            S_t &:= \sum_{i = 1}^{t} \delta_i \quad \forall t \in \N,
            \sigma_t &:= \sum_{i = 1}^{t}\tilde \eta_i \quad \forall t \in \N. 
        If there exists a sequence of $\eta_i, \tilde \eta_i$ as defined in
        \hyperref[def:ag_prox_grad_ppm]
        {definition \ref*{def:ag_prox_grad_ppm}} 
        such that
            & \Phi_{t + 1} - \Phi_{t} =
                \sum_{i = 1}^{t} \tilde \eta_i
            \right)\Upsilon_{2, t + 1}^{\text{AG}} 
            \le \delta_{t + 1} \quad 
            \forall t \in \mathbb N, 
            & \Upsilon_{1, 1}^{\text{AG}} \le \delta_1. 
        for some $\delta_i, i \in \N$. 
        Then 
            h(z_T) - h(x_*) &\le 
            \sigma_T^{-1}\left(
                S_{T} + \frac{1}{2}\Vert x_0 - x_*\Vert^2
            \right). 
        Where $x_*$ is an minimizer of $h$. 
        So $h(z_T) - h(x_*)$ has convergence rate $\mathcal O(\sigma_T^{-1})$ when $S_T \le 0$, and $\mathcal O(\sigma_T^{-1}S_T)$ when $S_T > 0$. 
        By definition we have
            \Phi_{t + 1} - \Phi_t 
                \sum_{i = 1}^{t+1} \tilde\eta_{i}
            \right) (h(z_{t + 1}) - h(x_*)) 
                \sum_{i = 1}^{t} \tilde\eta_{i}
            \right) (h(z_{t}) - h(x_*)) 
            + \frac{1}{2}\Vert x_t - x_*\Vert^2
            - \frac{1}{2}\Vert x_{t + 1} - x_*\Vert^2
            \tilde \eta_{t + 1} (h(z_{t + 1}) - h(z_*))
                \sum_{i = 1}^{t} \tilde \eta_i
            \right)(h(z_{t + 1}) - h(z_t))
            + \frac{1}{2}\Vert x_t - x_*\Vert^2
            - \frac{1}{2}\Vert x_{t + 1} - x_*\Vert^2
            &= \left(
                \sum_{i = 1}^{t} \tilde \eta_i
            \right)\Upsilon_{2, t + 1}^{\text{AG}} + \Upsilon_{1, t + 1}^{\text{AG}} \le \delta_{t + 1}. 
        Telescoping for $t = 0, \cdots, T- 1$
            \Phi_T - \Phi_0 = 
            \sum_{i = 0}^{T - 1}\Phi_{i + 1} - \Phi_i 
            \sum_{i = 0}^{T - 1}\delta_{i + 1}
            = S_{T}. 
        So then $\Phi_T - \Phi_0$ yields: 
            \sigma_T (h(z_T) - h(x_*)) 
            + \frac{1}{2}\Vert x_t - x_*\Vert^2 
            - \frac{1}{2}\Vert x_0 - x_*\Vert^2 
            &\le S_{T}
            \implies 
            \sigma_T(h(z_T) - h(x_*))
            S_T + \frac{1}{2}\Vert x_0 - x_*\Vert^2
            h(z_T) - h(x_*) &\le 
            \sigma_T^{-1}\left(
                S_{T} + \frac{1}{2}\Vert x_0 - x_*\Vert^2
        which yields a convergence rate $\mathcal O(\sigma_T^{-1}S_{T})$. 
        When $S_T = 0$, the convergence rate is $O(\sigma_T^{-1})$ instead. 

    \begin{theorem}[Constraints of PPM stepsize sequence]
    \label{thm:ag_generic_stepsize_constraints}\;\\
        Let iterates: $z_t, x_t, y_t$. 
        \hyperref[def:ag_prox_grad_generic]
        {definition \ref*{def:ag_prox_grad_generic}}. 
        If the stepsize $\eta_i, \tilde \eta_i$ satisfies relations 

@@

        Then for all $t \in \N$
            \epsilon_{t + 1} = \tilde \eta_{t + 1} - \eta_t - L^{-1}, 
        and 

@@

@@

\section{Recovering existing variants of Nesterov accelerated gradient}\label{sec:recovery}
    \subsection{Recovering FISTA}
        From 
        \hyperref[thm:ag_generic_stepsize_constraints]
        {theorem \ref*{thm:ag_generic_stepsize_constraints}}, 
        setting $\epsilon_i = 0$ for all $i$ yields $a_t = 1 + L\eta_t = \tilde \eta_{t + 1}$, for all $t \in \N$ where $a_t$ is the Nesterov sequence.
        In this case, there exists a momentum form of the generic form and it reduces to FISTA from Beck and Toubolle, or Similar Triangle Method as referred to by Ahn and Sra.

@@

            \hyperref[thm:ag_generic_stepsize_constraints]
                {theorem \ref*{thm:ag_generic_stepsize_constraints}}
